Transmembrane helices are also an overlooked source of major histocompatibility complex class II epitopes and evolutionary more conserved than expected by chance

### 2021-02-19 bbbq_article.tex
Selenoproteins to supmat

To determine the percentages of epitopes overlapping
with TMHs, we used reference proteomes of 
a human, viral and bacterial nature.
An overview of Uniprot IDs is shown in table \ref{tab:uniprot_ids}.
Instead of using the full human reference proteome, we
used the representative human reference proteome 
as used previously (\cite{bianchi2017}).

@@

The definition of a binder differs from \cite{bianchi2017}:
in our current study a peptide was called a binder if, within a certain haplotype, 
any of its 9-mer peptides have an IC50 value in the lowest 2\% of 
the epitopes within a 
\emph{proteome} (see tables \ref{tab:tmh_binders_mhc1} and \ref{tab:tmh_binders_mhc2}
for values), whereas the original study defined
a binder as having an IC50 in the lowest 2\% 
of the peptides within a \emph{protein}.
% See https://github.com/richelbilderbeek/bianchi_et_al_2017/blob/72e6755a31d400158368509fd80a41e984677ab1/predict-binders.R#L17
We believe our revised definition precludes bias of proteins 
that give rise to no or only very few MHC epitopes

@@

We recently showed that peptides derived from TMHs of human proteins 
were predicted to be over-presented in MHC-I. 
Here, we wondered if this finding would be general 
and if peptides stemming from TMHs of microbial pathogens 
would also be over-presented.
Figure \ref{fig:1} shows the percentages of MHC-I epitopes predicted to be overlapping 
with TMHs for our human, viral and bacterial proteome \richel{GvdB: you didn't explain yet which virus and bacteria}.
See the supplementary materials (table \ref{tab:tmh_binders_mhc1}) 
for the exact TMH and epitope counts.
\richel{GvdB: Also provide a short conclusion}

@@

\richel{GvdB: Sounds a but irrelevant to me. Do you really want to mention this in the discussion?}
% \paragraph{Removal of selenoproteins}
%
We have removed the 25 selenoproteins found in the human
reference proteome, as both prediction tools struggle with 
selenocysteine, the 21st amino acid incorporated in human proteins.
TMHMM gives a clear error message, forcing us to remove
these selenoproteins from the analysis.
How PureseqTM deals with selenocysteine is unclear,
as it gives no error, nor is it mentioned anywhere,
nor do the authors respond to emails. Due to this, we removed the
selenoproteins from our analysis.

@@

We thank the Center for Information Technology of the University 
of Groningen for its support and for providing access to the Peregrine 
high performance computing cluster. 
Additionally, we would like to thank Sci-Hub (\cite{himmelstein2018sci})
for allowing us to read paywalled articles while working from home.
\richel{GvdB: Should we include Maxim? He did the alignment of Myc tub. Let’s also ask Frans.}

@@

% \paragraph{TMH prediction software}

The earlier study uses TMHMM, which has a restricted software license.
Instead, we chose to use PureseqTM, which is free and open source.

@@

The earlier study uses \verb;epitope-prediction; a hand-crafted method, 
that has been trained on MHC-I haplotypes only,
which is used here again. For MHC-II IC50 predictions, the
'netmhc2pan' R package is used, which calls NetMHC2pan.
Both are packages are free and open-source.

@@

In this study, a peptide is defined a binder (for a haplotype), 
if \emp{within a proteome} in which it is found, 
is within the peptides with the 2\% lowest IC50 values.
Subsection \ref{subsec:ic50s_per_haplotype} shows the IC50 values
for a binder per haplotype.

% \paragraph{Unclear handing of sequences with U}

Within the human reference proteome, there are 30 sequences that have a
selenocysteine (with the symbol 'U').
See \url{https://github.com/richelbilderbeek/bbbq_article/issues/158}.

>>>>

We used a human, viral and bacterial reference proteome to 
determine the percentages of epitopes overlapping
with TMHs.
We used a human reference proteome
with UniProt ID UP000005640\_9606, which
is a representative set of the full proteome
and used is our previous study (\cite{bianchi2017}).
We used SARS-CoV-2 as our viral reference proteome
with UniProt ID UP000464024 and
Mycobacterium tuberculosis (MTb) 
with UniProt ID UP000001584.
We chose these pathogens based on their impact on the human population.

@@

The definition of a binder differs from \cite{bianchi2017}:
in our current study a peptide was called a binder if, within a certain haplotype, 
any of its 9-mer peptides have an IC50 value in the lowest 2\% of 
the epitopes within a 
\emph{proteome} (see tables \ref{tab:ic50_binders_mhc1} and \ref{tab:ic50_binders_mhc2}
for values), whereas the original study defined
a binder as having an IC50 in the lowest 2\% 
of the peptides within a \emph{protein}.
% See https://github.com/richelbilderbeek/bianchi_et_al_2017/blob/72e6755a31d400158368509fd80a41e984677ab1/predict-binders.R#L17
We believe our revised definition precludes bias of proteins 
that give rise to no or only very few MHC epitopes

@@

We recently showed that peptides derived from TMHs of human proteins 
were predicted to be over-presented in MHC-I. 
Here, we wondered if this finding would be general 
and if peptides stemming from TMHs of microbial pathogens 
would also be over-presented.
Figure \ref{fig:1} shows the percentages of MHC-I epitopes predicted to be overlapping 
with TMHs for our human, viral and bacterial proteome.
See the supplementary materials (table \ref{tab:tmh_binders_mhc1}) 
for the exact TMH and epitope counts.
\richel{GvdB: Also provide a short conclusion}

@@

@@

We thank the Center for Information Technology of the University 
of Groningen for its support and for providing access to the Peregrine 
high performance computing cluster. 
Additionally, we would like to thank Sci-Hub (\cite{himmelstein2018sci})
for allowing us to read paywalled articles while working from home.
\richel{GvdB: Should we include Maxim? He did the alignment of Mycobacterium tuberculosis. Let’s also ask Frans.}

@@

@@

The earlier study uses \verb;epitope-prediction; a hand-crafted method, 
that has been trained on MHC-I haplotypes only,
which is used here again. For MHC-II IC50 predictions, the
'mhcnuggetsr' R package is used.

@@

In this study, a peptide is defined a binder (for a haplotype), 
if within a \emp{proteome} in which it is found, 
is within the peptides with the 2\% lowest IC50 values,
where our previous study used the lowest 2\% IC50 values
within each \emp{protein}.
Subsection \ref{subsec:ic50s_per_haplotype} shows the IC50 values
for a binder per haplotype.

% \paragraph{Selenoproteins}

Our previous study used the TMHMM web server
to predict TMHs.
The desktop version of TMHMM, however, gives an
error message on the 25 selenoproteins found in the human
reference proteome.
For the sake of reproducible research, we use the desktop version (as
we can call it from scripts) and, due to this, we removed the
selenoproteins from this analysis.
